---
title: "UN Climate Report"
author: "Matthew Corp. Data Science Team"
date: '2022-06-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{r, echo = FALSE}
# Datasets
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(leaflet)
library(knitr)
library(pander)
library(ggthemes)

# Greenhouse gas emissions by country-year, 1750 - 2020
ghg <- read_csv('https://raw.githubusercontent.com/ericmkeen/capstone/master/co2.csv')

# Greenhouse gas emissions by country-year-sector, 1990 - 2018
sectors <- read_csv('https://raw.githubusercontent.com/ericmkeen/capstone/master/co2_sectors.csv')
```

# Introduction

Carbon Emissions have proved detrimental to the weather, wildlife, weather, and many aspects of this world. The following is a detailed report of Green House Gas (GHG) emissions, specially Carbon (Co2), over time across the globe.

 
The following report consist of data from two databases.

`GHG`: Detailed report of Country, Continent, and region Co2 emissions from 1750 to 2020

`sectors`: Detailed report of Co2 Emissions by Economic Sectors in Countries, Cotinents, and Regions around the globe.
 
  
  
Note: Carbon Emissions are measure in Parts Per Million (PPM)

# Present Day GHG Emissions

The following graphs will depict GHG emissions in Present Day on a geographic level. Note the years for each graph, as "Present Day" is defined on a different scale for one of them.

## Co2 Greatest Emmitters
Analysis of Top 10 Co2 Emitting Countries

```{r}
country <- sectors %>% 
  group_by(entity) %>% 
  filter(entity != "World" &
           entity != "High-income countries" &
           entity != "Upper-middle-income countries" &
           entity != "Lower-middle-income countries" &
           entity != "North America (excl. USA)" &
           entity != "Europe (excl. EU-28)" &
           entity != "Europe (excl. EU-27)" &
           entity != "Asia (excl. China & India)" &
           entity != "European Union (27)" &
           entity != "European Union (28)" &
           entity != "Europe" &
           entity != "Asia" &
           entity != "North America" &
           entity != "Africa" &
           entity != "Oceania" &
           entity != "Low-income countries") %>% 
  summarise(CO2 = sum(co2)) %>% 
  arrange(desc(CO2))

country_top10 <- sectors %>% 
  group_by(entity) %>% 
  filter(entity != "World" &
           entity != "High-income countries" &
           entity != "Upper-middle-income countries" &
           entity != "Lower-middle-income countries" &
           entity != "North America (excl. USA)" &
           entity != "Europe (excl. EU-28)" &
           entity != "Europe (excl. EU-27)" &
           entity != "Asia (excl. China & India)" &
           entity != "European Union (27)" &
           entity != "European Union (28)" &
           entity != "Europe" &
           entity != "Asia" &
           entity != "North America" &
           entity != "Africa" &
           entity != "Oceania" &
           entity != "Low-income countries") %>% 
  summarise(CO2 = sum(co2)) %>% 
  arrange(desc(CO2)) %>% 
  head(10)

country_bottom10 <- sectors %>% 
  group_by(entity) %>% 
  filter(entity != "World" &
           entity != "High-income countries" &
           entity != "Upper-middle-income countries" &
           entity != "Lower-middle-income countries" &
           entity != "North America (excl. USA)" &
           entity != "Europe (excl. EU-28)" &
           entity != "Europe (excl. EU-27)" &
           entity != "Asia (excl. China & India)" &
           entity != "European Union (27)" &
           entity != "European Union (28)" &
           entity != "Europe" &
           entity != "Asia" &
           entity != "North America" &
           entity != "Africa" &
           entity != "Oceania" &
           entity != "Low-income countries",
         co2 != 'NA') %>% 
  summarise(CO2 = sum(co2)) %>% 
  arrange(desc(CO2)) %>% 
  tail(10)
```

```{r}
ggplot(country_top10) +
  geom_histogram(aes(x = CO2, y = entity, fill = entity), stat = 'identity') +
  labs(subtitle = '1990 - 2018',
       caption = 'Ellie Davis',
       x = 'CO2 Levels',
       y= 'Country') +
  theme_classic() +
  theme(legend.position = 'none')
```

## Co2 Lowest Emmitters
Analysis of Bottom 10 Co2 Emitting Countries
```{r}
ggplot(country_bottom10) +
  geom_histogram(aes(x = CO2, y= entity, fill = entity), stat = 'identity') +
  labs(title = 'Least Co2 Emitters by Country in Recent Years',
       subtitle = '1990 - 2018',
       caption = 'Ellie Davis',
       x = 'CO2 Levels',
       y= 'Country') +
  theme_classic()+
   theme(legend.position = 'none') 
```
<!-- ## Recent Region Co2 Emissions -->
<!-- ```{r} -->
<!-- regions <- read_csv("C:\Users\jplus\Downloads\countriesOfWorld.csv") -->

<!-- regions <- regions %>%  -->
<!--   select(Country, Region) -->

<!-- regCO2 <- ghg %>%  -->
<!--   filter(Entity %in% regions$Country) %>%  -->
<!--   rename(Country = Entity) -->

<!-- regCO2 <- merge(regCO2, regions) %>%  -->
<!--   group_by(Region, Year) %>%  -->
<!--   summarize(co2 = sum(`Annual CO2 emissions (zero filled)`)) %>%  -->
<!--   mutate(co2 = co2/1000000) %>%  -->
<!--   filter(Year >= 1990 & Year <= 2018) -->

<!-- ggplot(data = regCO2, aes(x = Year, y = co2, color = Region)) + -->
<!--   geom_line() + -->
<!--   labs(title = "CO2 Emissions from 1990 to 2018", subtitle = "by Region", y = "CO2 Emissions (in Millions)") + -->
<!--   theme_classic() -->
<!-- ``` -->

## Recent Sector Co2 Emissions
Analysis of Co2 emissions by Sector from 1990 - 2018 
```{r}
test <- sectors %>% filter(entity != "World" &
           entity != "High-income countries" &
           entity != "Upper-middle-income countries" &
           entity != "Lower-middle-income countries" &
           entity != "North America (excl. USA)" &
           entity != "Europe (excl. EU-28)" &
           entity != "Europe (excl. EU-27)" &
           entity != "Asia (excl. China & India)" &
           entity != "European Union (27)" &
           entity != "European Union (28)" &
           entity != "Europe" &
           entity != "Asia" &
           entity != "North America" &
           entity != "Africa" &
           entity != "Oceania" &
           entity != "Low-income countries") %>% 
  group_by(sector) %>% 
  summarize(co2_e = sum(co2, na.rm = TRUE))

ggplot( data = test) + 
  geom_col(aes( x = sector, y = co2_e/1000000, fill = sector)) +
  labs( x = "Sector", y = "CO2 Emissions (Million PPM)", color = "Sector") + theme_clean() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  theme(legend.position = "None")

```





# Long Term Carbon Emissions
Ranging from 1750 - 2020, the following graphs will detail the long term Co2 emissions across multiple geographic locations.

## Global Long Term Carbon Emissions

```{r}
global <- ghg %>% 
  group_by(Year) %>%
  summarise(CO2 = sum(`Annual CO2 emissions (zero filled)`))
```

```{r}
ggplot(global, aes(x = Year, y = CO2/1000000)) +
  geom_line(color = 'dark green') +
  scale_x_continuous(breaks = seq(1750, 2020, by = 20)) +
  labs(title = 'CO2 Emissions Globally',
       subtitle = '1750 - 2020',
       caption = 'Ellie Davis',
       y = 'CO2 Emissions (Millions of PPM)')
```


## Co2 Emissions by Continent
```{r}
regions <- ghg %>% filter(Entity == "World" |
                            Entity == "High-income countries" |
                            Entity == "Upper-middle-income countries" |
                            Entity == "Lower-middle-income countries" |
                            Entity == "North America (excl. USA)" |
                            Entity == "Europe (excl. EU-28)" |
                            Entity == "Europe (excl. EU-27)" |
                            Entity == "Asia (excl. China & India)" |
                            Entity == "European Union (27)" |
                            Entity == "European Union (28)" |
                            Entity == "Europe" |
                            Entity == "Asia" |
                            Entity == "North America" |
                            Entity == "Africa" |
                            Entity == "Oceania" |
                            Entity == "Low-income countries")
ggplot(data = regions) + geom_line(aes(x = Year, y = regions$`Annual CO2 emissions (zero filled)`, color = Entity))

```

## Sector by Sector
Measuring Co2 Emissions by Economic Sector from 1990 - 2018
```{r}
#Creating DF for Sector by Sector Longterm CO2 Emissions
test <- sectors %>% filter(entity != "World" &
           entity != "High-income countries" &
           entity != "Upper-middle-income countries" &
           entity != "Lower-middle-income countries" &
           entity != "North America (excl. USA)" &
           entity != "Europe (excl. EU-28)" &
           entity != "Europe (excl. EU-27)" &
           entity != "Asia (excl. China & India)" &
           entity != "European Union (27)" &
           entity != "European Union (28)" &
           entity != "Europe" &
           entity != "Asia" &
           entity != "North America" &
           entity != "Africa" &
           entity != "Oceania" &
           entity != "Low-income countries") %>% 
  group_by(year, sector) %>% 
  summarize(co2_e = sum(co2, na.rm = TRUE))

ggplot( data = test) + 
  geom_line(aes( x = year, y = co2_e/1000000, color = sector)) + 
  geom_point(aes( x = year, y = co2_e/1000000, color = sector)) + 
  labs( x = "Year", y = "CO2 Emissions (Million PPM)", color = "Sector") + theme_clean()
```

## Income based Co2 Emissions
```{r}
region_by_income <- ghg %>% filter(Entity == "High-income countries" |
                                     Entity == "Upper-middle-income countries" |
                                     Entity == "Lower-middle-income countries" |
                                     Entity == "Low-income countries")
ggplot(data = region_by_income) +
  geom_line(aes(x = Year, y = region_by_income$`Annual CO2 emissions (zero filled)`, color = Entity)) +
  labs(title ="CO2 emissions by level of income from 1750 to 2020.", y = "Annual CO2 emissions")
```




# Trends in Top 5 Most Emitting Countries

## Long Term Co2 Emissions
```{r}
top5C <- ghg %>% 
  filter(Entity != "World" &
           Entity != "High-income countries" &
           Entity != "Upper-middle-income countries" &
           Entity != "Lower-middle-income countries" &
           Entity != "North America (excl. USA)" &
           Entity != "Europe (excl. EU-28)" &
           Entity != "Europe (excl. EU-27)" &
           Entity != "Asia (excl. China & India)" &
           Entity != "European Union (27)" &
           Entity != "European Union (28)" &
           Entity != "Europe" &
           Entity != "Asia" &
           Entity != "North America" &
           Entity != "Africa" &
           Entity != "Oceania" &
           Entity != "Low-income countries") %>% 
  group_by(Entity) %>% 
  summarize(totalCO2 = sum(`Annual CO2 emissions (zero filled)`)) %>% 
  arrange(desc(totalCO2)) %>% 
  head(5)

top5 <- ghg %>% 
  filter(Entity %in% top5C$Entity) %>% 
  mutate(`Annual CO2 emissions (zero filled)` = `Annual CO2 emissions (zero filled)`/1000000)

ggplot(data = top5, aes(x = Year, y = `Annual CO2 emissions (zero filled)`, color = Entity)) +
  geom_line() +
  labs(title = "Emissions of Top 5 Highest Emission Countries",
       subtitle = "1750 to 2020",
       y = "CO2 Emissions (in Millions)") +
  theme_clean() +
  scale_x_continuous(breaks = seq(1750, 2020, by = 25))
```

## Co2 by Sector
```{r}
top5_by_sector <- sectors %>% 
  filter(entity != "World" &
           entity != "High-income countries" &
           entity != "Upper-middle-income countries" &
           entity != "Lower-middle-income countries" &
           entity != "North America (excl. USA)" &
           entity != "Europe (excl. EU-28)" &
           entity != "Europe (excl. EU-27)" &
           entity != "Asia (excl. China & India)" &
           entity != "European Union (27)" &
           entity != "European Union (28)" &
           entity != "Europe" &
           entity != "Asia" &
           entity != "North America" &
           entity != "Africa" &
           entity != "Oceania" &
           entity != "Low-income countries",
         entity %in% c('United States', 'China', 'Russia', 'Germany', 'United Kingdom')) %>% 
  group_by(entity, sector) %>%
  summarize(totalCO2 = sum(co2))

ggplot(data = top5_by_sector, 
       aes(x = sector, y = totalCO2, fill = sector)) +
  geom_histogram(stat = 'identity') +
  labs(title = "Emissions of Top 5 Highest Emission Countries",
       subtitle = "1750 to 2020, by Sector",
       y = "CO2 Emissions (in Millions)",
       x = " ") +
  facet_wrap(~entity) +
  theme(axis.text.x = element_blank())
```

Data and code can be sent by request for further analysis. 

Collected and Analyzed by Matthew Corp.